The shape of dark matter halos in the AURIGA simulations

### 2019-03-05 Paper/paper.tex
edit

\section{Analysis}

@@

\citet{Deg_and_Widrow_2013} & $0.82$&$0.40$&
$20$kpc-$60$kpc$ $&$90^{\circ}$&
Mid-axis orientation. Sagittarius stream\\\hline \hline

@@

\begin{table*}
\begin{tabular}{|l|cc|cc|c|p{4cm}|}\hline
Reference&$q_{\rho}$&$s_{\rho}$&$q_{\phi}$&$s_{\phi}$&$R$& Methodology
\\ \hline \hline

\multirow{2}{*}{\citet{Abadi_et_al._2010}}&&&${0.98}$&${0.85}$&-&
Almost independent of radius. No feedback: boundary case\\\hline 
\end{tabular}
\caption{(TODO: compute analogues in isopotential or isodensity
  according to Binney and Tremaine) {\bf Al final estos numeros son
    isodensidad o isopotencial}}
\end{table*}

To compare our results against observational values have we to
translate first the isodensity into the corresponding isopotential
rations, in which observational constraints are usually presented.
For this purpose, we run a simple algorithm to find an
approximation of the shape of the isopotential contour.

@@

@@

As a consistency check we compare the isopotential results with the analytic expectation of the
$(1-q_{\phi})\approx \frac{1}{3}(1-q_{\rho})$
\citep{Binney_and_Tremaine_2008}, taking the volume-enclosed axial
ratios  as an approximation for the isodensity contour ratios
$q_{\rho}$. 
Although this analytic expression if meant to be used on the outer
regions of logarithmic axisymmetric halos, it works well as a first
approximation for nearly axisymmetric halos as those produced by our
simulations 
We find that the difference between the measured and the
approximated isopotential axial ratios is not bigger than $quantity
percent$ {\bf }.

Little discussion: Are our results congruent with observations? at
which radii, does the DM halo shape vary that much?. Which models are
favored?

We repeat this process for the different radii from table

@@

\subsection{The rounding effect of baryons}

From the previous characterization of radial shapes it is clear that
MHD halos are rounder than DM halos (i.e. axial ratios are bigger) at
every sampled radii. This can be compared on tables \ref{tabe:DM
  table} and \ref{tabe:MHD table} or at the representative example on
figure \ref{fig:DM_MHD}. From there, it is also noticeable that the
rounding effect of baryons is stronger at the disk regime, where the
DM halo is almost perfectly oblate. Furthermore, MHD halos tend
towards more oblate shapes (T < 0.5) despite DM halos tendency towards
more prolate shapes (T>0.5).

This rounding effect is expected from the gravitational effect of the
flattened axisymmetric galactic disk. It is also reasonable that this
effect is not as strong at $\approx 100$kpc, where the disk potential
is dimmer compared to the DM halo potential. Keeping this in mind, one
would expect that the rounding effect of baryons is related to some
galactic parameters such as its component masses and radii. However,
even for the parameter of highest correlation with this rounding
effect (the baryonic fraction), the relation is not clear nor
conclusive due to the dispersion from galaxy peculiarities. In figure
\ref{fig:Star_Density_effect} we plot the ratios $c/a$ compared to the
baryonic fraction of each galaxy. Although some linear tendency is
suggested qualitatively, the dispersion of the sample is very high to
obtain some conclusive relation. This is an evidence that
adiabatic-contraction models are not realistic as they may neglect
some effects of the galaxy evolution in the whole cosmological
context.

\textbf{Actually we have not examined the relation of c/a in MHD halos
  with some measure of c/a from the disk, that is something like
  Zdisk/Rdisk. This actually would make more sense from a physical
  point of view: effect of the potential.}

Although the effect of the baryonic disk on the shape of the DM halo
is a reasonable explanation for the rounding effect, it does not
actually explain the deviation from oblateness of MHD halos at
$r<10$kpc. In other words, if the disk is perfectly axisymmetric,
there must be some source of triaxiality at $r<10$kpc to explain the
low axial ratios.

\textit{ Talk about source of triaxiality at the inner parts of the
  halos (bar?). This source of triaxiality at the inner parts explains
  why the axial ratios are $\approx 0.95$ and not exactly $1$. We
  should also discuss that the decrease in the axial ratios for bigger
  radii may actually be bigger/steeper but it is dimmed by the
  contribution of inner parts.} 
%
%\begin{figure}
%  \centering
%  \subfloat[Level4 MHD Vs DM at inner regions]{\includegraphics[width=0.5\columnwidth]{./pics/Triaxial_Plane/Triaxiality_Inner_lvl4.png}}
%  \hfill
%  \subfloat[Level4 MHD Vs DM at outer regions]{\includegraphics[width=0.5\columnwidth]{./pics/Triaxial_Plane/Triaxiality_Outter_lvl4.png}}
%  \hfill
%  \caption{Axial ratios as shown on $c/a$ Vs $b/a$. Each dot represents a halo shape at some radius. Some observational constraints are plotted alongside our results. Here, dots are clustered, proving the general tendence of halos to get rounder on the outer parts.(Optimize space. caption replaces title.  Present constraints representation of density)}
%  \label{fig:Triaxiality_Inner_Outer}
%\end{figure}

@@

\section*{Acknowledgements}
This project has received funding from the European Union’s Horizon
2020 Research and Innovation Programme under the Marie
Sk\l{}odowska-Curie grant agreement No 734374.

>>>>

\subsection{What drives the rounding effect?}

From our characterization of radial shapes it is clear that
MHD halos are rounder than DM halos every sampled radii. 
It is also noticeable that the rounding effect of baryons is stronger
at the disk regime, where the DM halo is almost perfectly oblate. 
Furthermore, MHD halos tend towards more oblate shapes (T < 0.5)
despite DM halos tendency towards more prolate shapes (T>0.5). 
This rounding effect can be explained by gravitational effect of the
flattened axisymmetric galactic disk. 
It also explanes the weakness of this affect around $\approx 100$kpc,
where the disk potential is weaker compared to the DM halo potential.

{\bf Hace falta una grafica de la distribucion cumulativa de
  triaxialidad para (R\_vir y R\_vir/16 por separado), en cada panel se
  com para MHD vs DM.}

Following that linea reasoning, one would also expect that the rounding
effect of baryons is related to some  galactic parameters such as its
component masses and radii. 
We look for these kind of correlations and to find that the strongest
can be found for the stellar density.
However, the correlation is relatively low ({\bf cuanto vale el
  coeficiente de correlation}) most likely due to the different
formation histories.
In other words, the effect of the baryonic disk on the shape of the DM halo
does not fully explain the deviation from oblateness of MHD halos at
$r<10$kpc.

Figure  \ref{fig:Star_Density_effect} shows the
{\bf que es es lo que muestra esta grafica? falta las unidades de la
  densidad. y el eje vertical que es? es un cambio absoluto? fraccional?}

{\bf Finalmente cuales son las cantidades disponibles en el disco?
  cuales son las que muestran una mayor correlacion? Haria falta un
  plot con todas las correlaciones y la mencion de alguna prueba de
  machine learning del fit lineal para ver cuales son los parametros
  mas importantes.}

\textbf{Actually we have not examined the relation of c/a in MHD halos
  with some measure of c/a from the disk, that is something like
  Zdisk/Rdisk. This actually would make more sense from a physical
  point of view: effect of the potential. }

\textit{ Talk about source of triaxiality at the inner parts of the
  halos (bar?). This source of triaxiality at the inner parts explains
  why the axial ratios are $\approx 0.95$ and not exactly $1$. We
  should also discuss that the decrease in the axial ratios for bigger
  radii may actually be bigger/steeper but it is dimmed by the
  contribution of inner parts.}

@@

@@

\citet{Deg_and_Widrow_2013} & $0.82$&$0.40$&
$20$kpc $60$kpc$ $&$90^{\circ}$&
Mid-axis orientation. Sagittarius stream\\\hline \hline

Half of the observational constraints are computed in terms of
isodensity contours, the other half in terms of isopotential contours.
To compare our results against the second kind we must either
translate the isodensity results or recompute in terms of isopotential
regions.
For this purpose, we run a simple iterative algorithm to find an
approximation of the shape of the isopotential contour.

@@

In Figure \ref{fig:density_potential} we compare the two approaches to
measure the axial rations and plot the analytic expectation for $q$, 
$(1-q_{\phi})\approx \frac{1}{3}(1-q_{\rho})$
\citep{Binney_and_Tremaine_2008}, taking the volume-enclosed axial
ratios  as an approximation for the isodensity contour ratios
$q_{\rho}$. 
This plot is computed at four different radii.

Although the analytic expression is meant to be used on the outer
regions of logarithmic axisymmetric halos, it works well as a first
approximation for nearly axisymmetric halos as those produced by our
simulations 
We find that the difference between the measured and the
approximated isopotential axial ratios is not bigger than $quantity
percent$ {\bf }.

@@

@@

@@

\section*{Acknowledgements}
This project has received funding from the European Union's Horizon
2020 Research and Innovation Programme under the Marie
Sk\l{}odowska-Curie grant agreement No 734374.
